\vspace{-2.5em} Singular value decomposition of a dual orthogonal-view\\ polarized fluorescence microscope\vspace{-1em}

### 2018-07-12 notes/2018-05-30-polarized-dispim-svd/report/report.tex
Finish diSPIM 3D SVD notes.

In the
\href{https://github.com/talonchandler/polharmonic/blob/master/notes/2018-05-14-single-view-continuous-svd/report/report.pdf}{May
  14 notes} we derived the two-dimensional detection kernel. The kernel is
two-dimensional because it only considers thin in-focus objects. For a complete
non-paraxial three-dimensional model I would eventually like to follow Backer
and Moerner \cite{backer2014}, but there is still a few weeks of work required
to find the associated transfer functions. Instead I'll follow the early diSPIM
deconvolution models \cite{wu2013} and use an axial Gaussian function to model
the detection kernel.

The detection kernel for an objective along the $z$ axis (path A) as
\begin{align}
  h_{\tx{A}, \tx{det}}(\ro, \so) &= \left[[{a}^2(r_{xy}) + 2b^2(r_{xy})]y_0^0(\so) + \frac{1}{\sqrt{5}}\left[- a^2(r_{xy}) + 4b^2(r_{xy})\right]y_2^0(\so)\right]\mathfrak{g}(r_z, \sigma_{\tx{det}}), \label{eq:det1}
\end{align}
where
\begin{align}
  a(r) = \frac{J_1(2\pi \nu_or)}{\pi \nu_or}, 
  &\hspace{2em}
    b(r) = \frac{\tx{NA}}{n_o}\left[\frac{J_2(2\pi \nu_or)}{\pi \nu_or}\right],  \label{eq:abparadef}
    \intertext{and}
  r_{xy} \equiv \sqrt{r_x^2 + r_y^2},\hspace{2em}
  \nu_o \equiv &\frac{\tx{NA}}{\lambda},\hspace{2em}
  \tx{NA} = n_o\sin\alpha.
\end{align}

To find the two-dimensional detection kernel for path B we transform the
spherical harmonic coefficients using Eq. \ref{eq:transform}
\begin{align}
  h_{\tx{B}, \tx{det}}(\ro, \so) = \Bigg[[{a}^2(r_{yz}) + 2b^2(r_{yz})]y_0^0(\so) &+ \frac{1}{2\sqrt{5}}\left[- a^2(r_{yz}) + 4b^2(r_{yz})\right]y_2^0(\so) \nonumber \\ &+ \frac{1}{2}\sqrt{\frac{3}{5}}\left[- a^2(r_{yz}) + 4b^2(r_{yz})\right]y_2^2(\so)\Bigg]\mathfrak{g}(r_x, \sigma_{\tx{det}}), \label{eq:det2}
\end{align}
where $r_{yz} \equiv \sqrt{r_y^2 + r_z^2}$.

@@

\section{Transfer function}
We can rewrite Eqs. \ref{eq:fwd} and \ref{eq:adj} in the frequency domain as
\begin{align}
  G_{\tv, n}(\bv) &= \lmsum H_{l,n,\tv}^{m}(\bv)F_l^m(\bv), \label{eq:ffwd}\\
  F_l^m(\bv) &= \sum_{\tv \in \{\tx{A}, \tx{B}\}}\sum_{n=-\infty}^\infty H_{l,n,\tv}^{m}(\bv)G_{\tv, n}(\bv) \label{eq:fadj},
\end{align}
where
\begin{align}
  G_{\tv, n}(\bv) &= \intr{3}d\rd\, \me^{i2\pi\rd\cdot\bv}\ints{1}d\p\, z_n(\p) g_{\tv}(\rd, \p),\\
  H_{l,n,\tv}^m(\bv) &= \intr{3}d\ro\, \me^{i2\pi\ro\cdot\bv}\ints{1}d\p\, z_n(\p)\ints{2}d\so\, y_l^m(\so) h_{\tv}(\ro, \p, \so),\\
  F_l^m(\bv) &= \intr{3}d\ro\, \me^{i2\pi\ro\cdot\bv}\ints{2}d\so\, y_l^m(\so) f(\ro, \so).
\end{align}
In Figs. \ref{fig:tfA} and \ref{fig:tfB} we plot the transverse part of the
diSPIM transfer function. Extending these plots to three-dimensions is simple
given our assumption that the axial part is Gaussian. Notice that view B has a
non-zero element in the $m=-1$ column---this is a first for the microscopes
we've considered so far. Also notice that each view contributes new
information---we'll see this result in a clearer form when we find the complete
SVD.

\fig{../calculations/out/Hillx.pdf}{1.0}{Transverse transfer function
  $H_{l,n,\tx{A}}^m(\bv_{xy})$ for view A of a polarized diSPIM microscope with
  a 0.8 NA detection objective. \textbf{Rows:} $l$ indexes the object-space
  spherical harmonic band. \textbf{Columns:} $m$ indexes the object-space
  spherical harmonic degree, and $n$ indexes the data-space circular harmonic
  band. \textbf{Entries:} Each column and row contains a continuous
  two-dimensional plot indexed by the vector $\mb{r}_{xy}$ ranging from
  $-2\text{NA}/\lambda$ to $+2\text{NA}/\lambda$ (although $\mb{r}_{xy}$ is
  rotationally symmetric so one dimension would suffice). All values are
  normalized between $-1$ \textcolor{blue}{(blue)} and $+1$
  \textcolor{red}{(red)} with $0$ colored white.}{tfA}
  % filename, width, caption, label

\fig{../calculations/out/Hillz.pdf}{1.0}{Transverse transfer function
  $H_{l,n,\tx{B}}^m(\bv_{yz})$ for view B of a polarized diSPIM microscope with
  a 0.8 NA detection objective. See caption of Fig. \ref{fig:tfA} for
  details.}{tfB}
  % filename, width, caption, label

@@

Also notice that uniform distributions (or anything resembling a uniform
distribution) are not in the measurement space of either view. We saw the same
thing for single-view epi-illumination microscopes, although in that case there
were distributions that were mostly uniform with slight depressions along the
optical axis.

\fig{../calculations/out/SVSx.pdf}{1.0}{Singular system for view A
  ($x$-illumination, $z$-detection) of a polarized diSPIM microscope with a 0.8
  NA detection objective. \textbf{Rows:} Discrete branches of the singular value
  system indexed by $j$. \textbf{Column 1:} Continuous singular value spectra
  for each branch $j$ indexed by spatial frequency $\bs{\rho}$. The singular
  value spectra are normalized with contour lines at 0 and 0.1. \textbf{Columns
    2-5:} Angular part of the object-space singular vector at spatial
  frequencies marked with `x's in the first column. The camera is facing the
  origin along the [1,1,1] axis with the $z$ axis pointing up along the page. A
  \textcolor{red}{red} surface indicates positive values, a
  \textcolor{blue}{blue} surface indicates negative values, and the distance
  from the origin indicates the magnitude.}{sysA}
  % filename, width, caption, label

\fig{../calculations/out/SVSz.pdf}{1.0}{Singular system for view B
  ($z$-illumination, $x$-detection) of a polarized diSPIM microscope with a 0.8
  NA detection objective. See Fig. \ref{fig:sysA} caption for details. }{sysB}
  % filename, width, caption, label

\section{Dual-view singular value decomposition}
We will follow Burvall et al. \cite{burvall06} and the single-view case and
calculate the singular value decomposition in data space. We want to find the
eigenvalues $\mu_{\bp, j}$ and eigenfunctions $v_{\bp, j, \tv}(\rd, \p)$ that
satisfy
\begin{align}
  [\mc{H}\mc{H}^\dagger \mb{v}_{\bp, j}]_{\tv}(\rd, \p) = \mu_{\bp, j}v_{\bp, j, \tv}(\rd, \p). \label{eq:eig}
\end{align}
In the frequency domain the data-space singular functions will be in the form
\begin{align}
  v_{\bp, j, \tv}(\rd, \p) = \me^{i2\pi\bp\cdot\rd} \sum_{n=-\infty}^{\infty} \left[V_{n, \tv}(\bp)\right]_j z_n(\p). \label{eq:fdata}
\end{align}
Plugging Eqs. \ref{eq:ffwd}, \ref{eq:fadj}, and \ref{eq:fdata} in \ref{eq:eig} gives
\begin{align}
  \sum_{n=-\infty}^{n=\infty} \sum_{\tv \in \{\tx{A}, \tx{B}\}} \lmsum H_{l,\tv',n'}^m(\bp)H_{l,\tv,n}^m(\bp)\left[V_{n, \tv}(\bp)\right]_j = \mu_{\bp, j}\left[V_{n, \tv}(\bp)\right]_j. \label{eq:eig2}
\end{align}
If we combine the indices $n$ and $\tv$ into a multi-index $\mb{i} = (n, \tv)$ then we
can rewrite \ref{eq:eig2} as
\begin{align}
  \sum_{\mb{i}} \left[\lmsum H_{l,\mb{i}'}^m(\bp)H_{l,\mb{i}}^m(\bp)\right]\left[V_{\mb{i}}(\bp)\right]_j = \mu_{\bp, j}\left[V_{\mb{i}}(\bp)\right]_j. \label{eq:eig3}
\end{align}
Finally, we can rewrite Eq. \ref{eq:eig3} in matrix form as
\begin{align}
  \ms{K}(\bp)\ms{V}_j(\bp) = \mu_{\bp, j}\ms{V}_j(\bp),
\end{align}
where the entries of $\ms{K}(\bp)$ are given by
\begin{align}
  K_{\mb{i}, \mb{i}'} (\bp) = \lmsum H_{l,\mb{i}'}^m(\bp)H_{l,\mb{i}}^m(\bp). 
\end{align}
nFor the diSPIM, the matrix $\mathsf{K}$ has $6\times 6$ non-zero entries---two
views with three circular harmonics for each view---so solving this eigenvalue
problem at each point in frequency space is computationally feasible.

>>>>

In the
\href{https://github.com/talonchandler/polharmonic/blob/master/papers/basic-theory/paper/paper.pdf}{paper
  draft} we derive the two-dimensional detection kernel. Note that this model
differs from the
\href{https://github.com/talonchandler/polharmonic/blob/master/notes/2018-05-14-single-view-continuous-svd/report/report.pdf}{May
  14 model}---the old model contains an error that over predicts the
spatio-angular coupling by a factor of 4. The kernel is two-dimensional because
it only considers thin in-focus objects. For a complete non-paraxial
three-dimensional model I would eventually like to follow Backer and Moerner
\cite{backer2014}, but there is still a few weeks of work required to find the
associated transfer functions. Instead I'll follow the early diSPIM
deconvolution models \cite{wu2013} and use an axial Gaussian function to model
the detection kernel.

The detection kernel for an objective along the $z$ axis (path A) as
\begin{align}
  h_{\tx{A}, \tx{det}}(\ro, \so) = \left\{\left[a_1(r_{xy}) + \frac{\alpha^2}{4}a_2(r_{xy})\right]y_0^0(\so) + \frac{1}{\sqrt{5}}\left[-a_1(r_{xy}) + \frac{\alpha^2}{2}a_2(r_{xy})\right]y_2^0(\so)\right\}\mathfrak{g}(r_z, \sigma_{\tx{det}}), \label{eq:det1}
\end{align}
where
\begin{align}
  a_n(r) &\equiv \frac{n}{\pi}\left[\frac{J_n(2\pi\nu_o r)}{2\pi\nu_o r}\right],
           \intertext{and}
  r_{xy} \equiv \sqrt{r_x^2 + r_y^2},\hspace{2em}
  \nu_o &\equiv \frac{\tx{NA}}{\lambda},\hspace{2em}
  \alpha \equiv \frac{\tx{NA}}{n_0}.
\end{align}

To find the two-dimensional detection kernel for path B we transform the
spherical harmonic coefficients using Eq. \ref{eq:transform}
\begin{align}
  h_{\tx{B}, \tx{det}}(\ro, \so) = \Bigg\{\left[{a}_1(r_{yz}) + \frac{\alpha^2}{4}a_2(r_{yz})\right]y_0^0(\so) &+ \frac{1}{2\sqrt{5}}\left[- a_1(r_{yz}) + \frac{\alpha^2}{2}a_2(r_{yz})\right]y_2^0(\so) \nonumber \\ &+ \frac{1}{2}\sqrt{\frac{3}{5}}\left[- a_1(r_{yz}) + \frac{\alpha^2}{2}a_2(r_{yz})\right]y_2^2(\so)\Bigg\}\mathfrak{g}(r_x, \sigma_{\tx{det}}), \label{eq:det2}
\end{align}
where $r_{yz} \equiv \sqrt{r_y^2 + r_z^2}$.

@@

\section{Transfer function}
We can rewrite Eqs. \ref{eq:fwd} and \ref{eq:adj} in the frequency domain as
\begin{align}
  G_{\tv, n}(\bv) &= \lmsum H_{l,n,\tv}^{m}(\bv)F_l^m(\bv), \label{eq:ffwd}\\
  F_l^m(\bv) &= \sum_{\tv \in \{\tx{A}, \tx{B}\}}\sum_{n=-\infty}^\infty H_{l,n,\tv}^{m}(\bv)G_{\tv, n}(\bv) \label{eq:fadj},
\end{align}
where
\begin{align}
  G_{\tv, n}(\bv) &= \intr{3}d\rd\, \me^{i2\pi\rd\cdot\bv}\ints{1}d\p\, z_n(\p) g_{\tv}(\rd, \p),\\
  H_{l,n,\tv}^m(\bv) &= \intr{3}d\ro\, \me^{i2\pi\ro\cdot\bv}\ints{1}d\p\, z_n(\p)\ints{2}d\so\, y_l^m(\so) h_{\tv}(\ro, \p, \so),\\
  F_l^m(\bv) &= \intr{3}d\ro\, \me^{i2\pi\ro\cdot\bv}\ints{2}d\so\, y_l^m(\so) f(\ro, \so).
\end{align}
In Figs. \ref{fig:tfA} and \ref{fig:tfB} we plot the transverse part of the
diSPIM transfer function. Extending these plots to three dimensions is simple
given our assumption that the axial part is Gaussian. Notice that view B has a
non-zero element in the $m=-1$ column---this is a first for the microscopes
we've considered so far. Also notice that each view contributes new
information---we'll see this result in a clearer form when we find the complete
SVD.

\fig{../calculations/out2/Hillx.pdf}{1.0}{Transverse transfer function
  $H_{l,n,\tx{A}}^m(\bv_{xy})$ for view A of a polarized diSPIM microscope with
  a 0.8 NA detection objective. \textbf{Rows:} $l$ indexes the object-space
  spherical harmonic band. \textbf{Columns:} $m$ indexes the object-space
  spherical harmonic degree, and $n$ indexes the data-space circular harmonic
  band. \textbf{Entries:} Each column and row contains a continuous
  two-dimensional plot indexed by the vector $\bs{\nu}_{xy}$ ranging from
  $-2\text{NA}/\lambda$ to $+2\text{NA}/\lambda$ (although $\bs{\nu}_{xy}$ is
  rotationally symmetric so one dimension would suffice). All values are
  normalized between $-1$ \textcolor{blue}{(blue)} and $+1$
  \textcolor{red}{(red)} with $0$ colored white.}{tfA}
  % filename, width, caption, label

\fig{../calculations/out2/Hillz.pdf}{1.0}{Transverse transfer function
  $H_{l,n,\tx{B}}^m(\bv_{yz})$ for view B of a polarized diSPIM microscope with
  a 0.8 NA detection objective. See caption of Fig. \ref{fig:tfA} for
  details.}{tfB}
  % filename, width, caption, label

@@

Also notice that uniform distributions (or anything resembling a uniform
distribution) are not object-space singular functions of either view. We saw the
same thing for single-view epi-illumination microscopes, although in that case
there were object-space singular functions that were mostly uniform with slight
depressions along the optical axis. Note that we can still measure uniform
distributions with either view, but these distributions have a component in the
null space.

\fig{../calculations/out2/SVSx.pdf}{1.0}{Singular system for view A
  ($x$-illumination, $z$-detection) of a polarized diSPIM microscope with a 0.8
  NA detection objective. \textbf{Rows:} Discrete branches of the singular value
  system indexed by $j$. \textbf{Column 1:} Continuous singular value spectrum
  for each branch $j$ indexed by spatial frequency $\bs{\rho}$. The singular
  value spectra are normalized with contour lines at 0 and 0.1. \textbf{Columns
    2-5:} Angular part of the object-space singular vector at spatial
  frequencies marked with `x's in the first column. The camera is facing the
  origin along the [1,1,1] axis with the $z$ axis pointing up along the page. A
  \textcolor{red}{red} surface indicates positive values, a
  \textcolor{blue}{blue} surface indicates negative values, and the distance
  from the origin indicates the magnitude.}{sysA}
  % filename, width, caption, label

\fig{../calculations/out2/SVSz.pdf}{1.0}{Singular system for view B
  ($z$-illumination, $x$-detection) of a polarized diSPIM microscope with a 0.8
  NA detection objective. See Fig. \ref{fig:sysA} caption for details. }{sysB}
  % filename, width, caption, label

\section{Dual-view singular value decomposition}
We will follow Burvall et al. \cite{burvall06} and the single-view case and
calculate the singular value decomposition in data space. We want to find the
eigenvalues $\mu_{\bp, j}$ and eigenfunctions $v_{\bp, j, \tv}(\rd, \p)$ that
satisfy
\begin{align}
  [\mc{H}\mc{H}^\dagger \mb{v}_{\bp, j}]_{\tv}(\rd, \p) = \mu_{\bp, j}v_{\bp, j, \tv}(\rd, \p). \label{eq:eig}
\end{align}
In the frequency domain the data-space singular functions will be in the form
\begin{align}
  v_{\bp, j, \tv}(\rd, \p) = \me^{i2\pi\bp\cdot\rd} \sum_{n=-\infty}^{\infty} \left[V_{n, \tv}(\bp)\right]_j z_n(\p). \label{eq:fdata}
\end{align}
Plugging Eqs. \ref{eq:ffwd}, \ref{eq:fadj}, and \ref{eq:fdata} in \ref{eq:eig} gives
\begin{align}
  \sum_{n=-\infty}^{n=\infty} \sum_{\tv \in \{\tx{A}, \tx{B}\}} \lmsum H_{l,\tv',n'}^m(\bp)H_{l,\tv,n}^m(\bp)\left[V_{n, \tv}(\bp)\right]_j = \mu_{\bp, j}\left[V_{n, \tv}(\bp)\right]_j. \label{eq:eig2}
\end{align}
If we combine the indices $n$ and $\tv$ into a multi-index $\mb{i} = (n, \tv)$ then we
can rewrite \ref{eq:eig2} as
\begin{align}
  \sum_{\mb{i}} \left[\lmsum H_{l,\mb{i}'}^m(\bp)H_{l,\mb{i}}^m(\bp)\right]\left[V_{\mb{i}}(\bp)\right]_j = \mu_{\bp, j}\left[V_{\mb{i}}(\bp)\right]_j. \label{eq:eig3}
\end{align}
Finally, we can rewrite Eq. \ref{eq:eig3} in matrix form as
\begin{align}
  \ms{K}(\bp)\ms{V}_j(\bp) = \mu_{\bp, j}\ms{V}_j(\bp),
\end{align}
where the entries of $\ms{K}(\bp)$ are given by
\begin{align}
  K_{\mb{i}, \mb{i}'} (\bp) = \lmsum H_{l,\mb{i}'}^m(\bp)H_{l,\mb{i}}^m(\bp). 
\end{align}
For the diSPIM, the matrix $\mathsf{K}$ has $6\times 6$ non-zero entries---two
views with three circular harmonics for each view---so solving this eigenvalue
problem at each point in frequency space is computationally feasible.

Figures \ref{fig:sysx}--\ref{fig:sysz} show the singular systems for a symmetric
0.8 NA polarized illumination diSPIM microscope where the axial resolution
($\propto 1/\sigma_{\text{ax}}$) is 1/3 of the lateral resolution for each view.
I expect to modify this parameter based on bead measurements taken with the
final system.

\fig{../calculations/multi-out/SVS3Dx.pdf}{0.95}{Singular system for a symmetric
  0.8 NA polarized illumination diSPIM microscope. \textbf{Rows:} Discrete
  branches of the singular value system. \textbf{Column 1:} Maximum intensity
  projections of the normalized singular value spectrum for each branch indexed
  by 3D spatial frequency $\bs{\rho}$. \textbf{Columns 2-5:} Angular part of the
  object-space singular vector at spatial frequencies marked with `x's in the
  first column. The camera is facing the origin along the [1,1,1] axis with the
  $z$ axis pointing up along the page. A \textcolor{red}{red} surface indicates
  positive values, a \textcolor{blue}{blue} surface indicates negative values,
  and the distance from the origin indicates the magnitude.}{sysx}
  % filename, width, caption, label

\fig{../calculations/multi-out/SVS3Dy.pdf}{1.0}{Identical to Figure
  \ref{fig:sysx} with object space singular values along the $y$ axis.}{sysy}
  % filename, width, caption, label

\fig{../calculations/multi-out/SVS3Dz.pdf}{1.0}{Identical to Figure
  \ref{fig:sysx} with object space singular values along the $z$ axis.}{sysz}
  % filename, width, caption, label
